---
title: "Assignment_04"
author: "Megan Xiao"
date: "2025-09-16"
output: github_document
---

## Data Preparation and Cleaning

First, define the 5 vectors: Frequency, BloodPressure, FirstAssess, SecondAssess, and FinalDecision.

Frequency represents a vector of vital signs frequency, but it is unknown exactly what kind of vital sign it represents. The vector is standardized since the numbers go from 0 to 1.

BloodPressure represents a vector of blood pressure measurements ranging from 32 to 205. Lower values indicate hypotension (low blood pressure) and higher values indicate hypertension (high blood pressure).

FirstAssess represents a vector of the initial binary assessments of Frequency and BloodPressure. 1 equals a bad reading and 0 equals a good reading. SecondAssess and FinalDecision include a similar format, where 1 equals a high priority for medical intervention/bad reading and 0 equals a low priority/good reading.

```{r}
Frequency     <- c(0.6, 0.3, 0.4, 0.4, 0.2, 0.6, 0.3, 0.4, 0.9, 0.2)
BloodPressure <- c(103, 87, 32, 42, 59, 109, 78, 205, 135, 176)
FirstAssess   <- c(1, 1, 1, 1, 0, 0, 0, 0, NA, 1)    # bad=1, good=0
SecondAssess  <- c(0, 0, 1, 1, 0, 0, 1, 1, 1, 1)    # low=0, high=1
FinalDecision <- c(0, 1, 0, 1, 0, 1, 0, 1, 1, 1)    # low=0, high=1
```

```{r cars}

```

## Generate Basic Visualizations


## Interpretation and Discussion

## Submission