---
title: "Assignment_06"
author: "Megan Xiao"
date: "2025-10-06"
output: github_document
---

## Matrix Addition & Subtraction

Matrix operations combine matrices using addition, subtraction,
multiplication, and division, transforming large systems to model
real-world situations. Physics, engineering, and technology all require
matrices and matrix operations to represent various scenarios.<br> We
will first look at addition and subtraction. Let's create two 2x2
matrices:<br>

```{r}
A <- matrix(c(2, 0, 1, 3), ncol = 2)
B <- matrix(c(5, 2, 4, -1), ncol = 2)
```

First, let's add B to A.<br>

```{r}
A + B
```

Each value from B is added to each value from A, respectively, returning
a new 2x2 matrix. The order of the values doesn't matter either, as we
can see when we add A to B; this is due to the commutative property,
wherein the order of numbers in addition yields the same result.<br>

```{r}
B + A
```

Now, let's subtract B from A.<br>

```{r}
A - B
```

Each value from B is subtracted from each value from A, returning a new
2x2 matrix. When we switch the order of A and B, however, we get a
different result. Although the numbers are the same, their signs are
reversed.<br>

```{r}
B - A
```

## Create a Diagonal Matrix

A diagonal matrix consists of a square matrix where all values are zero
except for the principal diagonal elements, forming a "triangle" of
zeroes on the upper and lowers parts of the matrix. Diagonal matrices
are useful for many operations.<br> To build a diagonal matrix, we use
the diag() function. Let's create a 4x4 diagonal matrix, D:<br>

```{r}
D <- diag(c(4, 1, 2, 3))
D
```

For each field where the index \# of the row is not equal to the index
\# of the column in D[row, column], 0 must exist in that field. Scalar
matrices, identity matrices, and null matrices are examples of diagonal
matrices. <br> The number of numbers passed into diag() reflects the
number of rows and columns in the matrix. Since we passed four numbers
into diag() for D, we get a 4x4 matrix. Five numbers would yield a 5x5
matrix, six numbers would yield a 6x6 matrix, and so on.<br>

## Construct a Custom 5 x 5 Matrix

Let's try to create this matrix:<br>

```         
 [,1] [,2] [,3] [,4] [,5]
```

[1,] 3 1 1 1 1 [2,] 2 3 0 0 0 [3,] 2 0 3 0 0 [4,] 2 0 0 3 0 [5,] 2 0 0 0
3 <br> First, let's create a 5x5 diagonal matrix named "column" with the
principal elements being 3:<br>

```{r}
custom <- diag(c(3),ncol=5,nrow=5)
custom
```

Then, with indexing, we can select the first row and the second to fifth
columns of "custom," and fill those fields with the value 1. <br>

```{r}
custom[1,(2:5)] <- 1
custom
```

Using the same method, we can select the first column and the second to
fifth rows and fill those fields with the value 2.<br>

```{r}
custom[(2:5),1] <- 2
custom
```

We have obtained the desired matrix.<br>

## Summary

R provides powerful, built-in tools for matrices. Matrices can be
directly added to each other using basic operation symbols, and certain
matrices, like diagonal matrices, can be easily created with functions
like diag(). Additionally, subsetting pre-existing matrices allows us to
manipulate certain parts without changing the entire matrix.<br>
