---
title: "Module #2"
author: Megan Xiao
output: github_document
date: "2025-08-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Erroneous myMean Function

From Task 3, I create the vector "assignment2" and the function "myMean."
```{r}
# create numerical vector assignment2
assignment2 <- c(16, 18, 14, 22, 27, 17, 19, 17, 17, 22, 20, 22)

#create function myMean that returns the average of a vector
myMean <- function(assignment2) {
  return(sum(assignment) / length(someData))
}
```

Attempting to pass assignment2 into myMean results in the following error:

Error in myMean(assignment2) : object 'assignment' not found

```{r}
#myMean(assignment2) <- returns an error

```

The function contains two erroneous variables: (1) assignment and (2) someData. Data was not passed into any of these variables, nor were these variables initialized with any data. Therefore, R cannot run the myMean function.

## Correcting myMean

To correct myMean, the parameter name 'assignment' should replace 'assignment2' and 'someData' in the myMean function.

```{r}
#corrected myMean function
myMean <- function(assignment) {
  return(sum(assignment) / length(assignment))
}

#pass assignment2 into myMean
print(myMean(assignment2))

#verify myMean returns the correct result by comparing value with base R mean function
print(mean(assignment2) == myMean(assignment2))
```
