---
title: "Assignment_03"
author: "Megan Xiao"
date: "2025-09-12"
output: github_document
---

```{r}
library(ggplot2)
library(reshape)
library(dplyr)
```

## Define and Inspect Your Data

First, I create the following vectors "Name", "ABC_poll", and "CBS_poll."
"Name" contains a list of recent presidential candidates."ABC_poll" and "CBS_poll" represent the percentage (points) of respondents in favor of the corresponding candidate.


```{r}
Name <- c("Jeb", "Donald", "Ted", "Marco", "Carly", "Hillary", "Bernie")
ABC_poll <- c(4, 62, 51, 21, 2, 14, 15)
CBS_poll <- c( 12, 75, 43, 19, 1, 21, 19)
```

I combine these vectors into a DataFrame.

```{r}
df_polls <- data.frame(Name, ABC_poll, CBS_poll)
```

The resulting DataFrame has three columns and seven rows:

```{r}
df_polls
```
Let's look at the structure of the DataFrame.

```{r}
str(df_polls)
```
The "Name" column contains character values, while "ABC_poll" and "CBS_poll" contains numeric values. 

Head of df_polls:
```{r}
head(df_polls)
```
The head() function shows the first six rows of df_polls. Since df_polls only has seven rows, this function is not really necessary. It's better used for large data sets.

## Compute Summary Statistics

The mean, median, and range of the ABC polls:

```{r}
print(paste("Mean of ABC polls:",mean(df_polls$ABC_poll)))
print(paste("Median of ABC polls:",median(df_polls$ABC_poll)))
print(paste("Range of ABC polls:",max(df_polls$ABC_poll) - min(df_polls$ABC_poll)))

```
The mean, median, and range of the CBS polls:
```{r}
print(paste("Mean of CBS polls:",mean(df_polls$CBS_poll)))
print(paste("Median of CBS polls:",median(df_polls$CBS_poll)))
print(paste("Range of CBS polls:",max(df_polls$CBS_poll) - min(df_polls$CBS_poll)))
```
Add a column for the difference of percentages between CBS and ABC polls:

```{r}
df_polls$Diff <- abs(df_polls$CBS_poll - df_polls$ABC_poll)
df_polls
```

## Exploratory Data Analysis

Below, I create a bar plot that compares the percentages of respondents in favor of each candidate. The bars are ordered by the biggest difference in points between ABC polls and CBS polls.

```{r}
df_polls.melt <- melt(df_polls, id.vars = c("Name","Diff"),measure.vars = c("ABC_poll","CBS_poll"))

df_polls.melt %>%
  arrange(Diff) %>%
  mutate(Name = factor(Name, levels = unique(Name))) %>%
  ggplot(aes(x = Name, y = value, fill = factor(variable))) + 
  geom_bar(stat="identity", position = position_dodge()) +
  theme_minimal() +
  coord_flip() +
  labs(fill = "Source",
       y = "Percentage in-favor",
       x = "Name of Candidate",
       title = "Polling Points Ordered by Difference in Source")
```

## Discuss and Reflect

According to the data and the bar plot, Donald showed the largest discrepancy between ABC and CBS. Candidates showing larger discrepancies also had more respondents who were in favor of that candidate. The correlation between the number of respondents showing favor and the difference in the score may point to multicollinearity.
Using made-up data may introduce bias and inaccurate data, especially if one does not have any background knowledge in the subject. Made-up data is also not indicative of real life and, in this case, where the sample consists of little rows and columns, makes it challenging for researchers to use the data for training or testing. In real poll data, there would be other columns describing each candidate, such as those who were not in favor or those who were in favor who were Republican/Democrat. Extra columns would give more insight into how certain variables may be correlated and if they would have any effect on percentages in favor. 
