---
title: "Assignment_02"
author: "Megan Xiao"
date: "2025-09-12"
output: html_document
---

```{r}
library(ggplot2)
library(reshape)
```

## Define and Inspect Your Data

First, I create the following vectors "Name", "ABC_poll", and "CBS_poll."
"Name" contains a list of recent presidential candidates."ABC_poll" and "CBS_poll" represent the percentage (points) of respondents in favor of the corresponding candidate.


```{r}
Name <- c("Jeb", "Donald", "Ted", "Marco", "Carly", "Hillary", "Bernie")
ABC_poll <- c(4, 62, 51, 21, 2, 14, 15)
CBS_poll <- c( 12, 75, 43, 19, 1, 21, 19)
```

I combine these vectors into a DataFrame.

```{r}
df_polls <- data.frame(Name, ABC_poll, CBS_poll)
```

The resulting DataFrame has three columns and seven rows:

```{r}
df_polls
```
Let's look at the structure of the DataFrame.

```{r}
str(df_polls)
```
The "Name" column contains character values, while "ABC_poll" and "CBS_poll" contains numeric values. 

Head of df_polls:
```{r}
head(df_polls)
```
The head() function shows the first six rows of df_polls. Since df_polls only has seven rows, this function is not really necessary. It's better used for large data sets.

## Compute Summary Statistics

The mean, median, and range of the ABC polls:

```{r}
print(paste("Mean of ABC polls:",mean(df_polls$ABC_poll)))
print(paste("Median of ABC polls:",median(df_polls$ABC_poll)))
print(paste("Range of ABC polls:",max(df_polls$ABC_poll) - min(df_polls$ABC_poll)))

```
The mean, median, and range of the CBS polls:
```{r}
print(paste("Mean of CBS polls:",mean(df_polls$CBS_poll)))
print(paste("Median of CBS polls:",median(df_polls$CBS_poll)))
print(paste("Range of CBS polls:",max(df_polls$CBS_poll) - min(df_polls$CBS_poll)))
```
Add a column for the difference between CBS and ABC:

```{r}
df_polls$Diff <- abs(df_polls$CBS_poll - df_polls$ABC_poll)
df_polls
```

## Exploratory Data Analysis

```{r}
df_polls.melt <- melt(df_polls, id.vars = "Name", measure.vars = c("ABC_poll","CBS_poll"))
ggplot(df_polls.melt, aes(x = Name, y = variable, fill = factor(variable))) + geom_bar()
```