---
title: "Assignment_05"
author: "Megan Xiao"
date: "2025-09-29"
output: html_document
---

## Create the matrics
A matrix is a 2D array of numbers arranged in rows and columns. Mathematicians often use arrays to organize and work with structured information. Matrices are used in a wide variety of applications, such as physics, quantum mechanics, or computer graphic displays. <br>
R allows us to create matrices using the matrix() function. Here, we create two matrices: A and B. A consists of numbers 1 to 100 ascending top-down (column) and left-right (row). B consists of numbers 1 to 1000 ascending in a similar way to A. We specify the argument nrow = 10 to force both matrices to have only 10 rows. However, since ncol was not specified, R will create a new column and continue filling the matrix after the 10th row. <br>
```{r}
A <- matrix(1:100, nrow = 10)
B <- matrix(1:1000, nrow = 10)
```

## Inspect Dimensions
To inspect the number of rows and columns of our matrices, we can use the dim() function. This function returns two numbers; the first number is the number of rows in the matrix. The second is the number of columns. <br>

```{r}
dim(A)
```
<br>Matrix A has 10 rows and 10 columns. It is a square matrix because it has the same number of columns as it has rows. <br>
```{r}
dim(B)
```
<br>Matrix B has 10 rows and 100 columns. It is not a square matrix because it does not contain the same rows as it does columns.<br>

## Compute Inverse and Determinant
Now, we will find the inverse and determinant of matrices A and B. To find the inverse, the determinant of a matrix must not be 0.<br>
First, let's find the determinant of A:

```{r}
det(A)
```
The determinant of A is 0. Therefore, if we try to find the inverse of A, we get the following error:

```{r}
tryCatch(solve(A), error = function(e) e)
```
This is because our matrix is singular. Since the numbers are sequential, the rows are linearly dependent on one another, leading to a determinant of 0.
To find a non-zero determinant from a matrix, we would need to disrupt the sequential pattern in some way. We can do this by flattening the current matrix into a vector, shuffling the elements, and reshaping the shuffled elements back into a matrix.

```{r}
A_vec <- as.vector(A)
shuffled_A <- sample(A_vec)
A <- matrix(shuffled_A, nrow = 10)
A
```
Now, if when we use the det() function, we get a valid determinant:
```{r}
det(A)
```
This also allows us to find the inverse of A:
```{r}
solve(A)
```
Now, let's try to find the determinant of B.
```{r}
tryCatch(det(B), error = function(e) e)
```
Like matrix A, we get an error; however, this time, it's because B is not a square matrix. It has an unequal number of rows and columns. Therefore, we need to convert B into a square matrix.
We can indice B into the shape of a regular matrix, but this, just like A, will change the original values and meanings of B. 
```{r}
B_ind <- matrix(B[1:100], nrow = 10, ncol = 10)
B_ind
```
However, this does not solve all our problems. If we try to find the determinant of B_ind...

```{r}
det(B_ind)
```
We get 0. Like matrix A, we also need to shuffle the values out of order to get a non-zero determinant.
```{r}
B_ind.vec <- as.vector(B_ind)
B_ind <- sample(B_ind.vec)
B_ind <- matrix(B_ind, nrow = 10)
B_ind
```
Now, we will get a non-zero determinant:
```{r}
det(B_ind)
```
And if we use solve():
```{r}
solve(B_ind)
```
We get the inverse for each value in B.

## Multiplying Matrices
Another way we can get the inverse and determinants is by multiplying matrices. However, this creates a new, unique matrix that does not reflect the inverses/determinants of the original matrices.
To multiply
```{r}

```

## Summary

