---
title: "Assignment_05"
author: "Megan Xiao"
date: "2025-09-29"
output: html_document
---

## Create the matrics
A matrix is a 2D array of numbers arranged in rows and columns. Mathematicians often use arrays to organize and work with structured information. Matrices are used in a wide variety of applications, such as physics, quantum mechanics, or computer graphic displays.
R allows us to create matrices using the matrix() function. Here, we create two matrices: A and B. A consists of numbers 1 to 100 ascending top-down (column) and left-right (row). B consists of numbers 1 to 1000 ascending in a similar way to A. We specify the argument nrow = 10 to force both matrices to have only 10 rows. However, since ncol was not specified, R will create a new column and continue filling the matrix after the 10th row.
```{r}
A <- matrix(1:100, nrow = 10)
B <- matrix(1:1000, nrow = 10)
```

## Inspect Dimensions

```{r}
dim(A)
```

```{r}
dim(B)
```

## Compute Inverse and Determinant

```{r}
# For A
invA <- solve(A)
detA <- det(A)

# For B, use tryCatch to capture errors
invB <- tryCatch(solve(B), error = function(e) e)
detB <- tryCatch(det(B),   error = function(e) e)
```

## Summary

