---
title: "Assignment_12"
author: "Author: Megan Xiao\n"
date: "Date: 2025-11-10"
output: github_document
---

# Introduction
R Markdown is a file format that allows users to combine R code, visuals (such as plots and tables), and narrative text into a reproducible document. It is an extension of the Markdown language and supports executable code chunks.The main purpose of R is to enhance the legibility of commentary and code. R Markdown has many output formats, such as HTML, PDF, and Word.<br>
R Markdown has three main components: Headers for meta data, R code chunks, and narrative text. This paragraph, for example, is a narrative text. Below, we will also explore R code chunks and LaTeX (inline) expressions. <br> 
# LaTeX Math Expression
LaTeX is a typesetting system used for scientific and technical documents. It is mainly used to format complex mathematical equations, providing structure and consistency.
`$a^2 + b^2 = c^2$`

# Executable R Code

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
data("ChickWeight")

```

```{r}
ChickWeight_wide <- ChickWeight %>%
  group_by(Diet, Time) %>%
  summarise(mean_weight = mean(weight), .groups = "drop") 


ggplot(data = ChickWeight_wide, aes(x=Time, y = mean_weight, color= Diet)) + 
  geom_line() +
  labs(x = "Time (in days)",
       y= "Avg Weight",
       title = "Time vs Avg Weight of Chicks on Diff Diets") +
  theme_minimal()
```
# Reflection
Markdown is a very popular format for sharing reproducible code because it can integrate both text and visuals. Personally, I prefer Markdown over plain R files because they allow for greater understandability and ease of documentation. This was my first time using LaTeX, but I'm glad to have learned it because it will make integrating mathematical equations easier. I learned how to use inline LaTeX with dollar signs to display equations for cleaner presentations of formulas. <br>
Narrative text supports and describes code chunks, which are defined by three backticks and an R brace. The code chunks are where the actual data analysis takes place. Visuals and plots are then dynamically inserted into the final document. Chunks also have additional options, such as "echo=FALSE," to enable further customization.<br>
I did not face any particular challenge in this assignment. However, when I first began using R Markdown files, I did have trouble with text formatting and GitHub output. I learned that to put breaks after paragraphs, I needed to use the "<br>" tag, and that to output a markdown file to GitHub, I needed to specify "output: github_document" at the top of the document.<br>